---
title: "Fantasy Football Regression Analysis"
author: "2d Lt Lyssa White"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(caret)
library(ggplot2)
library(xlsx)
library(dplyr)
library(tidyr)
```

```{r}
#function to clean QB data
read_and_clean_sheet_QB <- function(file_path, sheet_name) {
  # Read the sheet WITHOUT column names so we can manually set them
  data <- read_excel(file_path, sheet = sheet_name, col_names = FALSE)
  
  # Set first row as column names
  col_names <- as.character(data[2, ])  # Convert second row to character column names
  data <- data[-c(1,2), ]  # Remove first and second row since it's now the column names
  
  # Define which columns to rename
  passing_cols <- 6:13   # Adjust these indices if needed
  rushing_cols <- 14:17  # Adjust these indices if needed
  
  # Rename passing columns
  col_names[passing_cols] <- paste0("Passing_", col_names[passing_cols])
  
  # Rename rushing columns
  col_names[rushing_cols] <- paste0("Rushing_", col_names[rushing_cols])
  
  # Assign new column names
  colnames(data) <- col_names
  
  # Ensure that the columns exist before removing them
  cols_to_remove <- c("TEAM", "POS", "FPTS")
  cols_to_remove <- cols_to_remove[cols_to_remove %in% colnames(data)]  # Keep only existing columns

  # Drop unwanted columns
  data <- data %>% dplyr::select(-all_of(cols_to_remove))
  
  # Convert numeric columns back (since they become characters after renaming)
  data <- type_convert(data)
  
  return(data)
}
```

```{r}
#function to clean data, sets correct col names
read_and_clean_sheet_RB <- function(file_path, sheet_name) {
  # Read the sheet WITHOUT column names so we can manually set them
  data <- read_excel(file_path, sheet = sheet_name, col_names = FALSE)
  
  # Set first row as column names
  col_names <- as.character(data[2, ])  # Convert second row to character column names
  data <- data[-c(1,2), ]  # Remove first and second row since it's now the column names
  
  # Define which columns to rename
  rushing_cols <- 6:9  # Adjust these indices if needed
  recieving_cols <- 10:13
  fumbles_cols <- 15
  
  # Rename rushing columns
  col_names[rushing_cols] <- paste0("Rushing_", col_names[rushing_cols])
  col_names[recieving_cols] <- paste0("Receiving_", col_names[recieving_cols])
  col_names[fumbles_cols] <- paste0("Fumbles_", col_names[fumbles_cols])
  
  # Assign new column names
  colnames(data) <- col_names
  
  # Ensure that the columns exist before removing them
  cols_to_remove <- c("TEAM", "POS", "FPTS")
  cols_to_remove <- cols_to_remove[cols_to_remove %in% colnames(data)]  # Keep only existing columns

  # Drop unwanted columns
  data <- data %>% dplyr::select(-all_of(cols_to_remove))
  
  # Convert numeric columns back (since they become characters after renaming)
  data <- type_convert(data)
  
  return(data)
}
```

```{r}
#function to clean data, sets correct col names
read_and_clean_sheet_WR <- function(file_path, sheet_name) {
  # Read the sheet WITHOUT column names so we can manually set them
  data <- read_excel(file_path, sheet = sheet_name, col_names = FALSE)
  
  # Set first row as column names
  col_names <- as.character(data[2, ])  # Convert second row to character column names
  data <- data[-c(1,2), ]  # Remove first and second row since it's now the column names
  
  # Define which columns to rename
  recieving_cols <- 6:13  # Adjust these indices if needed
  rushing_cols <- 14:17
  fumbles_cols <- 19
  
  # Rename columns
  col_names[recieving_cols] <- paste0("Receiving_", col_names[recieving_cols])
  col_names[rushing_cols] <- paste0("Rushing_", col_names[rushing_cols])
  col_names[fumbles_cols] <- paste0("Fumbles_", col_names[fumbles_cols])
  
  # Assign new column names
  colnames(data) <- col_names
  
  # Ensure that the columns exist before removing them
  cols_to_remove <- c("TEAM", "POS", "FPTS")
  cols_to_remove <- cols_to_remove[cols_to_remove %in% colnames(data)]  # Keep only existing columns

  # Drop unwanted columns
  data <- data %>% dplyr::select(-all_of(cols_to_remove))
  
  # Convert numeric columns back (since they become characters after renaming)
  data <- type_convert(data)
  
  return(data)
}
```

```{r}
#function to clean data, sets correct col names
read_and_clean_sheet_TE <- function(file_path, sheet_name) {
  # Read the sheet WITHOUT column names so we can manually set them
  data <- read_excel(file_path, sheet = sheet_name, col_names = FALSE)
  
  # Set first row as column names
  col_names <- as.character(data[2, ])  # Convert second row to character column names
  data <- data[-c(1,2), ]  # Remove first and second row since it's now the column names
  
  # Define which columns to rename
  recieving_cols <- 6:13  # Adjust these indices if needed
  rushing_cols <- 14:17
  fumbles_cols <- 19
  
  # Rename columns
  col_names[recieving_cols] <- paste0("Receiving_", col_names[recieving_cols])
  col_names[rushing_cols] <- paste0("Rushing_", col_names[rushing_cols])
  col_names[fumbles_cols] <- paste0("Fumbles_", col_names[fumbles_cols])
  
  # Assign new column names
  colnames(data) <- col_names
  
  # Ensure that the columns exist before removing them
  cols_to_remove <- c("TEAM", "POS", "FPTS")
  cols_to_remove <- cols_to_remove[cols_to_remove %in% colnames(data)]  # Keep only existing columns

  # Drop unwanted columns
  data <- data %>% dplyr::select(-all_of(cols_to_remove))
  
  # Convert numeric columns back (since they become characters after renaming)
  data <- type_convert(data)
  
  return(data)
}
```


```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
#load datasets
qb_file <- "ProjectDataWinter2025 (QB).xlsx"
rb_file <- "ProjectDataWinter2025 (RB).xlsx"
wr_file <- "ProjectDataWinter2025 (WR).xlsx"
te_file <- "ProjectDataWinter2025 (TE).xlsx"

qb_sheets <- excel_sheets(qb_file)[1:4]
rb_sheets <- excel_sheets(rb_file)[1:4]
wr_sheets <- excel_sheets(wr_file)[1:4]
te_sheets <- excel_sheets(te_file)[1:4]
```
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
#load datasets
qb_data_list <- lapply(qb_sheets, function(sheet) read_and_clean_sheet_QB(qb_file, sheet))
rb_data_list <- lapply(rb_sheets, function(sheet) read_and_clean_sheet_RB(rb_file, sheet))
wr_data_list <- lapply(wr_sheets, function(sheet) read_and_clean_sheet_WR(wr_file, sheet))
te_data_list <- lapply(te_sheets, function(sheet) read_and_clean_sheet_TE(te_file, sheet))
```
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
names(qb_data_list) <- qb_sheets
names(rb_data_list) <- rb_sheets
names(wr_data_list) <- wr_sheets
names(te_data_list) <- te_sheets
```

#check for missing values
I want all players to be included, so we'll impute missing values with averages

```{r}
check_null_values <- function(data, dataset_name) {
  # Count null (NA) values in each column
  na_counts <- colSums(is.na(data))
  
  # Print dataset name and the null counts per column
  cat("\n--- Missing Values Report for", dataset_name, "---\n")
  print(na_counts)
  cat("\n")
}
```

```{r}
# Check for null values in QB datasets
check_null_values(qb_data_list[[2]], "QB Data 2022")
check_null_values(qb_data_list[[3]], "QB Data 2023")
check_null_values(qb_data_list[[4]], "QB Data 2024")

# Check for null values in RB datasets
check_null_values(rb_data_list[[2]], "RB Data 2022")
check_null_values(rb_data_list[[3]], "RB Data 2023")
check_null_values(rb_data_list[[4]], "RB Data 2024")

# Check for null values in WR datasets
check_null_values(wr_data_list[[2]], "WR Data 2022")
check_null_values(wr_data_list[[3]], "WR Data 2023")
check_null_values(wr_data_list[[4]], "WR Data 2024")

# Check for null values in TE datasets
check_null_values(te_data_list[[2]], "TE Data 2022")
check_null_values(te_data_list[[3]], "TE Data 2023")
check_null_values(te_data_list[[4]], "TE Data 2024")

```
```{r}
count_total_missing <- function(data, dataset_name) {
  total_missing <- sum(is.na(data))  # Count total NAs in the dataset
  cat("\n--- Total Missing Values for", dataset_name, ": ", total_missing, "---\n")
  return(total_missing)  # Return the count if needed for further use
}
```
```{r}
# Apply function to all datasets
# QB Datasets
qb_missing_2022 <- count_total_missing(qb_data_list[[2]], "QB Data 2022")
qb_missing_2023 <- count_total_missing(qb_data_list[[3]], "QB Data 2023")
qb_missing_2024 <- count_total_missing(qb_data_list[[4]], "QB Data 2024")

# RB Datasets
rb_missing_2022 <- count_total_missing(rb_data_list[[2]], "RB Data 2022")
rb_missing_2023 <- count_total_missing(rb_data_list[[3]], "RB Data 2023")
rb_missing_2024 <- count_total_missing(rb_data_list[[4]], "RB Data 2024")

# WR Datasets
wr_missing_2022 <- count_total_missing(wr_data_list[[2]], "WR Data 2022")
wr_missing_2023 <- count_total_missing(wr_data_list[[3]], "WR Data 2023")
wr_missing_2024 <- count_total_missing(wr_data_list[[4]], "WR Data 2024")

# TE Datasets
te_missing_2022 <- count_total_missing(te_data_list[[2]], "TE Data 2022")
te_missing_2023 <- count_total_missing(te_data_list[[3]], "TE Data 2023")
te_missing_2024 <- count_total_missing(te_data_list[[4]], "TE Data 2024")

```
Some for QBs, 0s

No missing data for running backs

Since WR and TE rarely rush, the missing values represent 0 rushing attempts and 0 rushing yards, so we'll fill all these missing values with zeros

```{r}
# Fill missing rushing stats with 0
te_data_list <- lapply(te_data_list, function(df) {
  df %>%
    mutate(across(starts_with("Rushing_"), ~replace_na(.x, 0)))
})

wr_data_list <- lapply(wr_data_list, function(df) {
  df %>%
    mutate(across(starts_with("Rushing_"), ~replace_na(.x, 0)))
})

replace_missing_qb <- function(data) {
  data %>%
    mutate(across(starts_with("Passing_"), ~replace_na(.x, 0))) %>%
    mutate(across(starts_with("Rushing_"), ~replace_na(.x, 0)))
}

qb_data_list[[2]] <- replace_missing_qb(qb_data_list[[2]])  # QB Data 2022
qb_data_list[[3]] <- replace_missing_qb(qb_data_list[[3]])  # QB Data 2023
qb_data_list[[4]] <- replace_missing_qb(qb_data_list[[4]])  # QB Data 2024
```

```{r}
#convert everything to numeric except name
convert_numeric_except_name <- function(df) {
  df %>%
    mutate(across(-NAME, ~ as.numeric(.), .names = "converted_{.col}"))  # Convert everything except 'name' to numeric
}

qb_data_list <- lapply(qb_data_list, convert_numeric_except_name)
rb_data_list <- lapply(rb_data_list, convert_numeric_except_name)
wr_data_list <- lapply(wr_data_list, convert_numeric_except_name)
te_data_list <- lapply(te_data_list, convert_numeric_except_name)
```

```{r}
#clean names
clean_names <- function(df) {
  df %>%
    mutate(NAME = str_trim(NAME))  # Remove leading and trailing spaces
}

qb_data_list <- lapply(qb_data_list, clean_names)
rb_data_list <- lapply(rb_data_list, clean_names)
wr_data_list <- lapply(wr_data_list, clean_names)
te_data_list <- lapply(te_data_list, clean_names)

```

```{r}
#count outliers
count_outliers <- function(df, dataset_name) {
  # Identify numeric columns only
  numeric_cols <- df %>% dplyr::select(where(is.numeric))
  
  # Compute IQR for each column
  Q1 <- numeric_cols %>% summarise(across(everything(), ~ quantile(.x, 0.25, na.rm = TRUE)))
  Q3 <- numeric_cols %>% summarise(across(everything(), ~ quantile(.x, 0.75, na.rm = TRUE)))
  IQR_values <- Q3 - Q1
  
  # Compute lower and upper bounds
  lower_bound <- Q1 - 1.5 * IQR_values
  upper_bound <- Q3 + 1.5 * IQR_values
  
  # Count outliers for each column
  outlier_counts <- numeric_cols %>%
    summarise(across(everything(), ~ sum(.x < lower_bound[[cur_column()]] | .x > upper_bound[[cur_column()]], na.rm = TRUE)))
  
  # Sum total outliers across all numeric columns
  total_outliers <- sum(outlier_counts)
  
  # Return a data frame
  return(tibble(Dataset = dataset_name, Total_Outliers = total_outliers))
}

```

```{r}
# Count outliers in each dataset
outlier_summary <- bind_rows(
  count_outliers(qb_data_list[[2]], "QB Data 2022"),
  count_outliers(qb_data_list[[3]], "QB Data 2023"),
  count_outliers(qb_data_list[[4]], "QB Data 2024"),
  count_outliers(rb_data_list[[2]], "RB Data 2022"),
  count_outliers(rb_data_list[[3]], "RB Data 2023"),
  count_outliers(rb_data_list[[4]], "RB Data 2024"),
  count_outliers(wr_data_list[[2]], "WR Data 2022"),
  count_outliers(wr_data_list[[3]], "WR Data 2023"),
  count_outliers(wr_data_list[[4]], "WR Data 2024"),
  count_outliers(te_data_list[[2]], "TE Data 2022"),
  count_outliers(te_data_list[[3]], "TE Data 2023"),
  count_outliers(te_data_list[[4]], "TE Data 2024")
)

# Display summary table
library(knitr)
kable(outlier_summary, caption = "Total Outliers Per Dataset")

```
We have A LOT of outliers, we'll keep them for know because we know there wasn't errors in collecting the data, if they impact our model negatively we will address them later. We could potentially want the outliers because we want to pick the best players and our assumption is that the best players are above average.

```{r}
qb_data_2022 <- qb_data_list[[2]]
qb_data_2023 <- qb_data_list[[3]]
qb_data_2024 <- qb_data_list[[4]]

rb_data_2022 <- rb_data_list[[2]]
rb_data_2023 <- rb_data_list[[3]]
rb_data_2024 <- rb_data_list[[4]]

wr_data_2022 <- wr_data_list[[2]]
wr_data_2023 <- wr_data_list[[3]]
wr_data_2024 <- wr_data_list[[4]]

te_data_2022 <- te_data_list[[2]]
te_data_2023 <- te_data_list[[3]]
te_data_2024 <- te_data_list[[4]]
```
```{r}
library(dplyr)

merge_data_2022 <- function(qb_data_2022, qb_data_2023) {
  # Rename fpts/g and FPTS/G columns in 2023 dataset
  qb_data_2023 <- qb_data_2023 %>%
    rename(`2023_fpts/g` = `FPTS/G`)  # Handles case sensitivity
  
  # Merge with 2022 dataset on "NAME"
  qb_combined <- qb_data_2022 %>%
    left_join(qb_data_2023 %>% dplyr::select(NAME, `2023_fpts/g`), by = "NAME")
  
  return(qb_combined)
}

qb_combined_2022 <- merge_data_2022(qb_data_2022, qb_data_2023)
rb_combined_2022 <- merge_data_2022(rb_data_2022, rb_data_2023)
wr_combined_2022 <- merge_data_2022(wr_data_2022, wr_data_2023)
te_combined_2022 <- merge_data_2022(te_data_2022, te_data_2023)
```

```{r}
#rows where 2023 fpts is NA
na_rows_qb <- qb_combined_2022 %>%
  filter(is.na(`2023_fpts/g`)) %>%
  dplyr::select(NAME)

print(na_rows_qb)
```
we have 18 QB rows with NAs. We'll go a google search and see if we can get their statistics, but if they didn't play in 2023, then we'll delete them because their information is not helpful to us. None of these players played in 2023, so we'll delete them from our data set.

```{r}
#rows where 2023 fpts is NA
na_rows_rb <- rb_combined_2022 %>%
  filter(is.na(`2023_fpts/g`)) %>%
  dplyr::select(NAME)

print(na_rows_rb)
```
```{r}
#rows where 2023 fpts is NA
na_rows_wr <- wr_combined_2022 %>%
  filter(is.na(`2023_fpts/g`)) %>%
  dplyr::select(NAME)

print(na_rows_wr)
```
```{r}
#rows where 2023 fpts is NA
na_rows_te <- te_combined_2022 %>%
  filter(is.na(`2023_fpts/g`)) %>%
  dplyr::select(NAME)

print(na_rows_te)
```
```{r}
#delete players who did not play in 2023
remove_na_players <- function(qb_combined_2022, na_rows_qb) {
  # Remove rows where NAME exists in na_rows_qb
  qb_filtered <- qb_combined_2022 %>%
    filter(!NAME %in% na_rows_qb$NAME)
  
  return(qb_filtered)
}

qb_data_2022 <- remove_na_players(qb_combined_2022, na_rows_qb)
rb_data_2022 <- remove_na_players(rb_combined_2022, na_rows_rb)
wr_data_2022 <- remove_na_players(wr_combined_2022, na_rows_wr)
te_data_2022 <- remove_na_players(te_combined_2022, na_rows_te)
```
Now again for 2023 and 2024
```{r}
library(dplyr)

merge_data_2023 <- function(qb_data_2023, qb_data_2024) {
  # Rename fpts/g and FPTS/G columns in 2023 dataset
  qb_data_2024 <- qb_data_2024 %>%
    rename(`2024_fpts/g` = `FPTS/G`)  # Handles case sensitivity
  
  # Merge with 2022 dataset on "NAME"
  qb_combined <- qb_data_2023 %>%
    left_join(qb_data_2024 %>% dplyr::select(NAME, `2024_fpts/g`), by = "NAME")
  
  return(qb_combined)
}

qb_combined_2023 <- merge_data_2023(qb_data_2023, qb_data_2024)
rb_combined_2023 <- merge_data_2023(rb_data_2023, rb_data_2024)
wr_combined_2023 <- merge_data_2023(wr_data_2023, wr_data_2024)
te_combined_2023 <- merge_data_2023(te_data_2023, te_data_2024)
```

```{r}
#rows where 2023 fpts is NA
na_rows_qb <- qb_combined_2023 %>%
  filter(is.na(`2024_fpts/g`)) %>%
  dplyr::select(NAME)

print(na_rows_qb)
```
we have 21 QB rows with NAs. We'll go a google search and see if we can get their statistics, but if they didn't play in 2023, then we'll delete them because their information is not helpful to us. None of these players played in 2023, so we'll delete them from our data set.

```{r}
#rows where 2023 fpts is NA
na_rows_rb <- rb_combined_2023 %>%
  filter(is.na(`2024_fpts/g`)) %>%
  dplyr::select(NAME)

print(na_rows_rb)
```
```{r}
#rows where 2023 fpts is NA
na_rows_wr <- wr_combined_2023 %>%
  filter(is.na(`2024_fpts/g`)) %>%
  dplyr::select(NAME)

print(na_rows_wr)
```
```{r}
#rows where 2023 fpts is NA
na_rows_te <- te_combined_2023 %>%
  filter(is.na(`2024_fpts/g`)) %>%
  dplyr::select(NAME)

print(na_rows_te)
```
```{r}
#delete players who did not play in 2023
remove_na_players <- function(qb_combined_2022, na_rows_qb) {
  # Remove rows where NAME exists in na_rows_qb
  qb_filtered <- qb_combined_2022 %>%
    filter(!NAME %in% na_rows_qb$NAME)
  
  return(qb_filtered)
}

qb_data_2023 <- remove_na_players(qb_combined_2023, na_rows_qb)
rb_data_2023 <- remove_na_players(rb_combined_2023, na_rows_rb)
wr_data_2023 <- remove_na_players(wr_combined_2023, na_rows_wr)
te_data_2023 <- remove_na_players(te_combined_2023, na_rows_te)
```

```{r}
colnames(qb_data_2022) <- tolower(colnames(qb_data_2022))
colnames(qb_data_2023) <- tolower(colnames(qb_data_2023))
colnames(qb_data_2024) <- tolower(colnames(qb_data_2024))

colnames(rb_data_2022) <- tolower(colnames(rb_data_2022))
colnames(rb_data_2023) <- tolower(colnames(rb_data_2023))
colnames(rb_data_2024) <- tolower(colnames(rb_data_2024))

colnames(wr_data_2022) <- tolower(colnames(wr_data_2022))
colnames(wr_data_2023) <- tolower(colnames(wr_data_2023))
colnames(wr_data_2024) <- tolower(colnames(wr_data_2024))

colnames(te_data_2022) <- tolower(colnames(te_data_2022))
colnames(te_data_2023) <- tolower(colnames(te_data_2023))
colnames(te_data_2024) <- tolower(colnames(te_data_2024))
```

```{r}
remove_converted_columns <- function(df) {
  df %>%
    dplyr::select(-starts_with("converted_"))  # Remove all columns that start with "converted_"
}

# Apply to QB datasets
qb_data_2022 <- remove_converted_columns(qb_data_2022)
qb_data_2023 <- remove_converted_columns(qb_data_2023)
qb_data_2024 <- remove_converted_columns(qb_data_2024)

# Apply to RB datasets
rb_data_2022 <- remove_converted_columns(rb_data_2022)
rb_data_2023 <- remove_converted_columns(rb_data_2023)
rb_data_2024 <- remove_converted_columns(rb_data_2024)

# Apply to WR datasets
wr_data_2022 <- remove_converted_columns(wr_data_2022)
wr_data_2023 <- remove_converted_columns(wr_data_2023)
wr_data_2024 <- remove_converted_columns(wr_data_2024)

# Apply to TE datasets
te_data_2022 <- remove_converted_columns(te_data_2022)
te_data_2023 <- remove_converted_columns(te_data_2023)
te_data_2024 <- remove_converted_columns(te_data_2024)
```

```{r}
#extract needed data
library(readr)

# Save QB data
write_csv(qb_data_2022, "qb_data_2022.csv")
write_csv(qb_data_2023, "qb_data_2023.csv")
write_csv(qb_data_2024, "qb_data_2024.csv")

# Save RB data
write_csv(rb_data_2022, "rb_data_2022.csv")
write_csv(rb_data_2023, "rb_data_2023.csv")
write_csv(rb_data_2024, "rb_data_2024.csv")

# Save WR data
write_csv(wr_data_2022, "wr_data_2022.csv")
write_csv(wr_data_2023, "wr_data_2023.csv")
write_csv(wr_data_2024, "wr_data_2024.csv")

# Save TE data
write_csv(te_data_2022, "te_data_2022.csv")
write_csv(te_data_2023, "te_data_2023.csv")
write_csv(te_data_2024, "te_data_2024.csv")
```





