---
title: "FF_FinalPredictions"
author: "2d Lt Lyssa White"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(caret)
library(ggplot2)
library(xlsx)
library(dplyr)
library(tidyr)
library(car)        # For VIF (Variance Inflation Factor)
library(leaps)      # For stepwise selection
library(MASS)       # For stepwise AIC/BIC
library(caret) 
library(knitr)
library(kableExtra)
library(corrplot)
library(reshape2)
library(MuMIn)
```

```{r}
#read in data
qb_predictions <- read.csv("qb_predictions.csv")
rb_predictions <- read.csv("rb_predictions.csv")
wr_predictions <- read.csv("wr_predictions.csv")
te_predictions <- read.csv("te_predictions.csv")
```

Let's merge them all and see what top 20 looks like
```{r}
predictions2024 <- rbind(qb_predictions, rb_predictions, wr_predictions, te_predictions)
sorted_preds <- predictions2024 %>% arrange(desc(predictions2024$predicted_fpts_2024))
```

```{r}
head(sorted_preds, 25)
```
We don't see a ton of QBs in top 20, usually their higher scoring so this tells us our model is under-estimating points for QBs but that's okay. We can look at the QB data separately to determine which to pick when we get to that point, but QB wasn't going to be our first or second choice anyway.

First choice: CeeDee Lamb
20th choice (pending 2-19 were taken): Evan Engram

# overall accuracy


```{r}
#scatterplot of actual vs predicted fantasy points
ggplot(sorted_preds, aes(x = `X2024_fpts.g`, y = `predicted_fpts_2024`, color = position)) +
  geom_point(size = 3, alpha = 0.8) +  # Scatter points colored by position
  geom_smooth(method = "lm", se = FALSE) +  # Separate regression lines per position
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +  # Perfect prediction line
  labs(title = "Predicted vs. Actual Fantasy Points per Game",
       x = "Actual FPPG",
       y = "Predicted FPPG",
       color = "Position") +  # Legend title
  theme_minimal() +
  theme(legend.position = "right")
```
```{r}
sorted_actual <- predictions2024 %>% filter(predictions2024$position != "QB") %>% arrange(desc(`X2024_fpts.g`))
head(sorted_actual, 50)
```

```{r}
top_20_preds <- tolower(trimws(sorted_preds$name[1:20]))  
top_20_actual <- tolower(trimws(sorted_actual$name[1:20]))

matching_players <- intersect(top_20_preds, top_20_actual)
print(matching_players)
```
```{r}
# Count the number of correct predictions within ±2 points
correct_predictions <- sum(abs(predictions2024$`predicted_fpts_2024` - predictions2024$`X2024_fpts.g`) <= 3)

# Calculate accuracy as a percentage
accuracy <- (correct_predictions / nrow(predictions2024)) * 100

# Print the accuracy
print(paste("Accuracy within ±3 fantasy points:", round(accuracy, 2), "%"))
```
```{r}
library(dplyr)

# Calculate accuracy for each position
position_accuracy <- predictions2024 %>%
  group_by(position) %>%
  summarise(
    total_players = n(),  # Count players per position
    correct_predictions = sum(abs(`predicted_fpts_2024` - `X2024_fpts.g`) <= 3Indiv),  # Count correct predictions
    accuracy = (correct_predictions / total_players) * 100  # Compute accuracy percentage
  )

# Print the accuracy table
print(position_accuracy)
```



