---
title: "FF_ModelBuilding"
author: "2d Lt Lyssa White"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(caret)
library(ggplot2)
library(xlsx)
library(dplyr)
library(tidyr)
library(car)        # For VIF (Variance Inflation Factor)
library(leaps)      # For stepwise selection
library(MASS)       # For stepwise AIC/BIC
library(caret) 
library(knitr)
library(kableExtra)
library(corrplot)
library(reshape2)
library(MuMIn)
```

```{r}
rb_data_2022 <- read.csv("rb_data_2022.csv")
rb_data_2023 <- read.csv("rb_data_2023.csv")
```

```{r}
rb_data_2022 <- rb_data_2022 %>%
    rename(`2023_fpts/g` = `X2023_fpts.g`)
rb_data_2023 <- rb_data_2023 %>%
    rename(`2024_fpts/g` = `X2024_fpts.g`)
```

Model Building

Since we know that rank is directly correlated with the average number of fantasy points per season for each of our players, we'll remove rank from the data. If we leave it in there could be high multicollinearity between rank and fantasy points per season which could result in unstable coefficients, inflated variance of coefficients, and decreased statistical significance. We'll choose to keep average fantasy points per season for the previous season here because this predictor is already standardized and is more specific than rank. We expect there to be multicollinearity present between other variables as well, since we know some of our metrics are calculated using the metrics provided. For example, in the quarterback dataset, passing completion percentage is calculated using passes completed and passes attempted, which are both also variables in the dataset. Since it isn't clear which variables will be better predictors, this concern will be addressed with a multicollinearity assessment later on. Name was also removed at this point to maintain only numeric regressors for the linear regression model.

```{r}
#remove rank and name from 2022 data
rb_data_2022 <- rb_data_2022 %>% dplyr::select(-c("rk"))
```
# Initial Model & All Possible Regressions

We'll use all data to run the all possible regressions, and then use K fold cross validation on the top candidate models.
```{r}
standardize_data <- function(df) {
  numeric_cols <- df %>% dplyr::select(where(is.numeric)) %>% names()
  
  means <- df %>% summarise(across(all_of(numeric_cols), mean, na.rm = TRUE)) 
  sds <- df %>% summarise(across(all_of(numeric_cols), sd, na.rm = TRUE))  
  
  scaled_df <- df %>%
    mutate(across(all_of(numeric_cols), ~ (. - means[[cur_column()]]) / sds[[cur_column()]]))
  
  attr(scaled_df, "means") <- means
  attr(scaled_df, "sds") <- sds
  
  return(scaled_df)
}

rb_data_2022 <- standardize_data(rb_data_2022)
```


Run RB full model and analyze p values to determine which variables to remove

```{r}
rb_full_model <- lm(`2023_fpts/g` ~ gp +rushing_att+rushing_yds+rushing_avg+rushing_td+receiving_tgts+receiving_rec+receiving_yds+receiving_td+fum+fumbles_lost+fpts.g, data = rb_data_2022, na.action="na.fail")
summary(rb_full_model)
```
```{r}
coef(rb_full_model)
```

check normality
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
rstudent_residuals <- rstudent(rb_full_model)
fitted_values <- fitted(rb_full_model)
```
```{r, echo=FALSE}
qqnorm(rstudent_residuals, main = "RB Normal Probability Plot")
qqline(rstudent_residuals, col="red", lwd=2)
```
We see that our residuals are mostly aligned with our reference line. This indicates that our residuals are distributed approximately normal. Next, let's look at a residual plot.

```{r, echo=FALSE}
plot(fitted_values, rstudent_residuals, xlab = "Fitted Values", ylab = "Residuals", main="RB Fitted Values vs R Student Residuals")
abline(h=0, col="red", lty=2)
```
We see from this residual plot that there is a random scattering of our residuals around the zero line, and they seem to be contained within a horizontal band with the exception of two outliers. This tells us that our variance of errors is constant and our data appears to be linear. Now let's look at the residuals vs each of our regressor values.

```{r, echo=FALSE}
selected_regressors <- c("gp", "rushing_att", "rushing_yds", "rushing_avg", "rushing_td", "receiving_tgts", "receiving_rec", "receiving_yds", "receiving_td", "fum", "fumbles_lost", "fpts.g")

for (regressor in selected_regressors) {
  plot(rb_data_2022[[regressor]], rstudent_residuals,
       main = paste("R-Student Residuals vs ", regressor),
       xlab = regressor,
       ylab = "Residuals")
  abline(h=0, col="red", lty=2)
}
```
mostly random scatterings, same outlier impacting all

These random scatterings confirm our assumption that the relationship between the selected regressors and $y$ is respectively linear. 

Influential Obs
Next we'll look at any pure outliers, leverage points, and influential points in our data. We'll evaluate hat values, Cook's D values, DFBETA values, DFFITS values and COVRATIO values.

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
hat_values = hatvalues(rb_full_model)
cooks_d = cooks.distance(rb_full_model)
dfbetas_values <- as.data.frame(dfbetas(rb_full_model))
dffits = dffits(rb_full_model)
covratio = covratio(rb_full_model)
```
```{r, echo=FALSE, warning=FALSE}
results_df <- data.frame(
  "RB" = rb_data_2022$name,
  "Studentized Residuals" = rstudent_residuals,
  "Hat Values" = hat_values,
  "Cooks D Values" = cooks_d,
  "DFBETA Values" = dfbetas_values,
  "DFFITS Values" = dffits,
  "COVRATIO Values" = covratio
)

results_df <- results_df %>%
  mutate(across(where(is.numeric), round, digits = 4))


kable(results_df, caption = "Regression Diagnostics Table", align="c", booktabs = TRUE) %>%
  kable_styling(latex_options = c("scale_down", "hold_position", "centering"))
```

Now let's identify what those thresholds are for our model.
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
p <- 13
n <- nrow(rb_data_2022)

hat <- (2*p)/n
dffits_val <- 2*sqrt(p/n)
dfbetas_vals <- 2/sqrt(n)
covratio_val1 <- 1 - (3*p)/n
covratio_val2 <- 1 + (3*p)/n
```

```{r, echo=FALSE}
influence_metrics2 <- data.frame(
  Metric = c("Studentized Residuals (R-Student)", 
             "Hat Values", 
             "Cook’s Distance", 
             "DFFITS", 
             "DFBETAs", 
             "COVRATIO"),

  Threshold = c(
    ">|2|, >|3|",
    round(hat, 4), 
    "> 1", 
    round(dffits_val, 4),
    round(dfbetas_vals, 4),
    paste("<", round(covratio_val1,4), "or >", round(covratio_val2,4))
  )
)

kable(influence_metrics2, caption = "Influence Metrics Thresholds Based on NFL Data") 
```
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
results_df$`Studentized.Residuals` <- as.numeric(results_df$`Studentized.Residuals`)
results_df$`Hat.Values` <- as.numeric(results_df$`Hat.Values`)
results_df$`Cooks.D.Values` <- as.numeric(results_df$`Cooks.D.Values`)
results_df$`DFFITS.Values` <- as.numeric(results_df$`DFFITS.Values`)
results_df$`DFBETA.Values.gp` <- as.numeric(results_df$`DFBETA.Values.gp`)
results_df$`DFBETA.Values.receiving_tgts` <- as.numeric(results_df$`DFBETA.Values.receiving_tgts`)
results_df$`DFBETA.Values.receiving_rec` <- as.numeric(results_df$`DFBETA.Values.receiving_rec`)
results_df$`DFBETA.Values.receiving_yds` <- as.numeric(results_df$`DFBETA.Values.receiving_yds`)
results_df$`DFBETA.Values.receiving_td` <- as.numeric(results_df$`DFBETA.Values.receiving_td`)
results_df$`DFBETA.Values.fum` <- as.numeric(results_df$`DFBETA.Values.fum`)
results_df$`DFBETA.Values.fumbles_lost` <- as.numeric(results_df$`DFBETA.Values.fumbles_lost`)
results_df$`DFBETA.Values.rushing_att` <- as.numeric(results_df$`DFBETA.Values.rushing_att`)
results_df$`DFBETA.Values.rushing_yds` <- as.numeric(results_df$`DFBETA.Values.rushing_yds`)
results_df$`DFBETA.Values.rushing_avg` <- as.numeric(results_df$`DFBETA.Values.rushing_avg`)
results_df$`DFBETA.Values.rushing_td` <- as.numeric(results_df$`DFBETA.Values.rushing_td`)
results_df$`DFBETA.Values.fpts.g` <- as.numeric(results_df$`DFBETA.Values.fpts.g`)
results_df$`COVRATIO.Values` <- as.numeric(results_df$`COVRATIO.Values`)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
outlier_rows <- list(
  "Studentized Residuals" = which(abs(results_df$`Studentized.Residuals`) > 3),
  "Hat Values" = which(results_df$`Hat.Values` > round(hat, 4)),
  "Cook’s Distance" = which(results_df$`Cooks.D.Values` > 1),
  "DFFITS" = which(abs(results_df$`DFFITS.Values`) > round(dffits_val, 4)),
  "DFBETAs" = which(abs(results_df$`DFBETA.Values.gp`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.receiving_tgts`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.receiving_rec`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.receiving_yds`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.receiving_td`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.fum`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.fumbles_lost`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.rushing_att`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.rushing_yds`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.rushing_avg`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.rushing_td`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.fpts.g`) > round(dfbetas_vals, 4)),
  "COVRATIO" = which(results_df$`COVRATIO.Values` < round(covratio_val1,4) | 
                     results_df$`COVRATIO.Values` > round(covratio_val2,4))
)

# Display rows for each metric
print(outlier_rows)
```
Since DFBETAS and COVRATIO don't tell us much on their own, we'll only look at rows that have at least one other irregular flag.

```{r}
studentized_residuals <- c(63)
hat_values <- c(17,41,50,66)
cooks_distance <- integer(0)  # Empty
dffits <- c(6,7,13,41,63,66)

dfbetas <- c(6,7,10,13,14,16,21,22,23,24,28,36,38,41,45,48,50,63,64,65,66)

covratio <- c(2,9,10,12,17,19,26,41,50,61,63,64,66)
```
```{r}
all_indices <- sort(unique(c(studentized_residuals, hat_values, dffits)))
dfbetas_filtered <- dfbetas[dfbetas %in% all_indices]
covratio_filtered <- covratio[covratio %in% all_indices]
```
```{r}
df <- data.frame(
  Index = all_indices,
  Studentized_Residuals = ifelse(all_indices %in% studentized_residuals, "✓", ""),
  Hat_Values = ifelse(all_indices %in% hat_values, "✓", ""),
  DFFITS = ifelse(all_indices %in% dffits, "✓", ""),
  DFBETAs = ifelse(all_indices %in% dfbetas_filtered, "✓", ""),
  COVRATIO = ifelse(all_indices %in% covratio_filtered, "✓", "")
)

print(df)
```
```{r}
df %>%
  kable(format = "pipe", align = "c") %>%
  kable_styling(full_width = FALSE)
```
We have 8 flagged points all together.
4 appearing in 3 or more categories

Indices with high studentized residuals have unusually large errors compared to the rest of the model. Indices with irregular hat values have high leverage and could influence the regression fit. Indices with irregular DFFITS values are influential points that might change the model significantly. Indices with irregular DFBETAS values cause bias in parameter estimated. Indices with irregular COVRATIO values may increase the standard errors of the coefficients.

We'll identify the most problmatic observations as those that appear in at least three categories. Since these are genuine observations and not data entry errors, we're going to choose to keep them. They represent real variation in the dataset, and removing them may limit the generalizability of the model. In addition, players that represent outliers on the higher end, may be exactly what we're looking for. Our goal is prediction and these outliers represent real world cases so we'll keep them. They may aslo be the result of regressors in the model that don't end up being important, or, they may be exacerbated by the regressors that are inportant. We'll keep them in mind and ensure to check them in our final models.

multicollinearity
check correlation matrix for pairwise collinearity
```{r, echo=FALSE}
cor_matrix <- cor(rb_data_2022[, selected_regressors], use = "pairwise.complete.obs")
#print(cor_matrix)
kable(cor_matrix, digits = 2, format = "markdown", caption="NFL Data Correlation Matrix")
```
```{r, echo=FALSE}
cor_melt <- melt(cor_matrix)

high_corr <- cor_melt %>%
  filter(abs(value) > 0.75 & value < 1) %>%
  arrange(desc(value))

high_corr <- high_corr %>%
  rowwise() %>%
  mutate(pair = paste0(sort(c(Var1, Var2)), collapse = "-")) %>%  # Sort pairs
  distinct(pair, .keep_all = TRUE) %>%
  dplyr::select(-pair)

kable(high_corr, digits = 2, format = "markdown", caption = "Highly Correlated Regressor Pairs")
```
10 highly correlated pairs, let's see if we can identify groups that are highly collinear.

```{r, echo=FALSE}
eigenvalues <- eigen(cor_matrix)$values
kable(data.frame(Eigenvalues = round(eigenvalues, 4)), format = "markdown", caption="Eigenvalues for NFL Data")
```
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
eigen_decomp <- eigen(cor_matrix)
eigenvectors <- eigen_decomp$vectors
smallest_eigenvector <- eigenvectors[, which.min(eigen_decomp$values)]

print(smallest_eigenvector)
```

```{r}
vif(rb_full_model)
```
6 above 10

```{r}
compute_press <- function(model) {
  residuals <- residuals(model)
  hat_values <- hatvalues(model)
  
  # Avoid division by zero for leverage = 1
  hat_values[hat_values == 1] <- 0.9999
  
  press_residuals <- sum((residuals / (1 - hat_values))^2, na.rm = TRUE)  # PRESS
  return(press_residuals)
}
```

```{r}
combinations <- dredge(rb_full_model, extra = c(R_Sq = function(x) summary(x)$r.squared,
R_Sq_Adj = function(x) summary(x)$adj.r.squared, MS_Res = function(x) summary(x)$sigma^2, Cp,
MallowCp = function(x) summary(x)$sigma^2*df.residual(x)/summary(rb_full_model)$sigma^2
-dim(rb_data_2022)[1]+2*length(x$coefficients), PRESS=compute_press))
```

```{r}
kable(
  combinations %>%
    head(10) %>%  # Select first 10 rows
    mutate(across(where(is.numeric), ~ round(.x, 3))),  # Round numeric columns
  format = "markdown",
  digits = 3,
  caption = "All Possible Regressions"
) %>%
  kable_styling(full_width = FALSE)  # Ensure it fits on the page
```
```{r}
# Add a column to store original model numbers
combinations <- combinations %>%
  mutate(Model_Number = rownames(combinations)) 
combinations %>% head(10)
```



```{r}
# Select top 10 models while keeping model number
clean_combinations <- combinations %>%
  head(20) %>%  # Select first 10 rows
  mutate(across(where(is.numeric), ~ round(.x, 3)))  # Round numeric columns

# Sort models by Mallows Cp and adjusted R^2 while retaining original model numbers
table_mallowscp <- combinations %>%
  arrange(MallowCp)

table_adj_r2 <- combinations %>%
  arrange(desc(R_Sq_Adj))

table_press <- combinations[order(combinations$PRESS), ] 

# Display Top 10 Models Sorted by Mallows Cp
kable(
  table_mallowscp %>%
    head(10) %>%
    mutate(across(where(is.numeric), ~ round(.x, 3))),  # Round to 3 decimals
  format = "markdown",
  digits = 3,
  caption = "Top 10 Models Sorted by Mallows Cp Statistic"
) %>%
  kable_styling(full_width = FALSE)

# Display Top 10 Models Sorted by Adjusted R^2
kable(
  table_adj_r2 %>%
    head(10) %>%
    mutate(across(where(is.numeric), ~ round(.x, 3))),  # Round to 3 decimals
  format = "markdown",
  digits = 3,
  caption = "Top 10 Models Sorted by Adjusted R^2"
) %>%
  kable_styling(full_width = FALSE)

# Display Top 10 Models Sorted by Adjusted R^2
kable(
  table_press %>%
    head(10) %>%
    mutate(across(where(is.numeric), ~ round(.x, 3))),  # Round to 3 decimals
  format = "markdown",
  digits = 3,
  caption = "Top 10 Models Sorted by PRESS"
) %>%
  kable_styling(full_width = FALSE)
```

3074, 1066, 1282, 3234 have both the best Mallow's CP and PRESS values. 1258 has the highest adjusted R^2 so well look that too.

```{r}
model_A <- lm(`2023_fpts/g` ~ fpts.g + rushing_td + rushing_yds, data = rb_data_2022)
model_B <- lm(`2023_fpts/g` ~ fpts.g + gp + receiving_td + rushing_td, data = rb_data_2022)
model_C <- lm(`2023_fpts/g` ~ fpts.g + rushing_att + rushing_td, data = rb_data_2022)
model_D <- lm(`2023_fpts/g` ~ fpts.g + receiving_td + receiving_yds +rushing_td + rushing_yds, data = rb_data_2022)
model_E <- lm(`2023_fpts/g` ~ fpts.g + gp + receiving_td + receiving_tgts + receiving_yds + rushing_td, data=rb_data_2022)
```

# Model Adequacy
summaries

what's strongly significant, what's borderline, stronger reasons now for including the terms; Mallow's CP, Adj R^2, take p values with a grain of salt
```{r}
sum_A <- summary(model_A)
sum_B <- summary(model_B)
sum_C <- summary(model_C)
sum_D <- summary(model_D)
sum_E <- summary(model_E)
sum_A
```
fpts.g is strongly significantand rushing_td is significant, rushing_yds is borderline significant

```{r}
sum_B
```
fpts.g and rushing_td are strongly significant. gp and receiving_td are borderline significant but again thats okay, still overall significant.

```{r}
sum_C
```
```{r}
sum_D
```
```{r}
sum_E
```


VIFs: center maybe? best regressors in spite of collinearity
```{r}
vifs_A <- vif(model_A)
vifs_B <- vif(model_B)
vifs_C <- vif(model_C)
vifs_D <- vif(model_D)
vifs_E <- vif(model_E)
vifs_A
vifs_B
vifs_C
vifs_D
vifs_E
```
models D and E may have issues with collinearity

Normal PP and residual plots
Model A
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
rstudent_residuals_A <- rstudent(model_A)
fitted_values_A <- fitted(model_A)
```
```{r, echo=FALSE}
qqnorm(rstudent_residuals_A, main = "Model A: RB Normal Probability Plot")
qqline(rstudent_residuals_A, col="red", lwd=2)
```
```{r, echo=FALSE}
plot(fitted_values_A, rstudent_residuals_A, xlab = "Fitted Values", ylab = "Residuals", main="Model A: RB Fitted Values vs R Student Residuals")
abline(h=0, col="red", lty=2)
```
```{r, echo=FALSE}
selected_regressors <- c("rushing_td", "rushing_yds", "fpts.g")

for (regressor in selected_regressors) {
  plot(rb_data_2022[[regressor]], rstudent_residuals_A,
       main = paste("Model A: R-Student Residuals vs ", regressor),
       xlab = regressor,
       ylab = "Residuals")
  abline(h=0, col="red", lty=2)
}
```

Model B
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
rstudent_residuals_B <- rstudent(model_B)
fitted_values_B <- fitted(model_B)
```
```{r, echo=FALSE}
qqnorm(rstudent_residuals_B, main = "Model B: RB Normal Probability Plot")
qqline(rstudent_residuals_B, col="red", lwd=2)
```
```{r, echo=FALSE}
plot(fitted_values_B, rstudent_residuals_B, xlab = "Fitted Values", ylab = "Residuals", main="Model B: RB Fitted Values vs R Student Residuals")
abline(h=0, col="red", lty=2)
```
```{r, echo=FALSE}
selected_regressors <- c("fpts.g", "gp", "receiving_td", "rushing_td")

for (regressor in selected_regressors) {
  plot(rb_data_2022[[regressor]], rstudent_residuals_B,
       main = paste("Model B: R-Student Residuals vs ", regressor),
       xlab = regressor,
       ylab = "Residuals")
  abline(h=0, col="red", lty=2)
}
```
Model C
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
rstudent_residuals_C <- rstudent(model_C)
fitted_values_C <- fitted(model_C)
```
```{r, echo=FALSE}
qqnorm(rstudent_residuals_C, main = "Model C: RB Normal Probability Plot")
qqline(rstudent_residuals_C, col="red", lwd=2)
```
```{r, echo=FALSE}
plot(fitted_values_C, rstudent_residuals_C, xlab = "Fitted Values", ylab = "Residuals", main="Model C: RB Fitted Values vs R Student Residuals")
abline(h=0, col="red", lty=2)
```
```{r, echo=FALSE}
selected_regressors <- c("fpts.g", "rushing_att", "rushing_td")

for (regressor in selected_regressors) {
  plot(rb_data_2022[[regressor]], rstudent_residuals_C,
       main = paste("Model C: R-Student Residuals vs ", regressor),
       xlab = regressor,
       ylab = "Residuals")
  abline(h=0, col="red", lty=2)
}
```
Model D
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
rstudent_residuals_D <- rstudent(model_D)
fitted_values_D <- fitted(model_D)
```
```{r, echo=FALSE}
qqnorm(rstudent_residuals_D, main = "Model D: RB Normal Probability Plot")
qqline(rstudent_residuals_D, col="red", lwd=2)
```
```{r, echo=FALSE}
plot(fitted_values_D, rstudent_residuals_D, xlab = "Fitted Values", ylab = "Residuals", main="Model D: RB Fitted Values vs R Student Residuals")
abline(h=0, col="red", lty=2)
```
```{r, echo=FALSE}
selected_regressors <- c("fpts.g", "receiving_td", "receiving_yds", "rushing_td", "rushing_yds")

for (regressor in selected_regressors) {
  plot(rb_data_2022[[regressor]], rstudent_residuals_D,
       main = paste("Model D: R-Student Residuals vs ", regressor),
       xlab = regressor,
       ylab = "Residuals")
  abline(h=0, col="red", lty=2)
}
```
Model E
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
rstudent_residuals_E <- rstudent(model_E)
fitted_values_E <- fitted(model_E)
```
```{r, echo=FALSE}
qqnorm(rstudent_residuals_E, main = "Model E: RB Normal Probability Plot")
qqline(rstudent_residuals_E, col="red", lwd=2)
```
```{r, echo=FALSE}
plot(fitted_values_E, rstudent_residuals_E, xlab = "Fitted Values", ylab = "Residuals", main="Model E: RB Fitted Values vs R Student Residuals")
abline(h=0, col="red", lty=2)
```
```{r, echo=FALSE}
selected_regressors <- c("gp", "fpts.g", "receiving_td", "receiving_tgts", "receiving_yds", "rushing_td")

for (regressor in selected_regressors) {
  plot(rb_data_2022[[regressor]], rstudent_residuals_E,
       main = paste("Model E: R-Student Residuals vs ", regressor),
       xlab = regressor,
       ylab = "Residuals")
  abline(h=0, col="red", lty=2)
}
```
problem children - all models are wrong, some models are useful
Model A
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
rstudent_residuals = rstudent(model_A)
hat_values = hatvalues(model_A)
cooks_d = cooks.distance(model_A)
dfbetas_values <- as.data.frame(dfbetas(model_A))
dffits = dffits(model_A)
covratio = covratio(model_A)
```
```{r, echo=FALSE, warning=FALSE}
results_df <- data.frame(
  "RB" = rb_data_2022$name,
  "Studentized Residuals" = rstudent_residuals,
  "Hat Values" = hat_values,
  "Cooks D Values" = cooks_d,
  "DFBETA Values" = dfbetas_values,
  "DFFITS Values" = dffits,
  "COVRATIO Values" = covratio
)

results_df <- results_df %>%
  mutate(across(where(is.numeric), round, digits = 4))


kable(results_df, caption = "Model A: Regression Diagnostics Table", align="c", booktabs = TRUE) %>%
  kable_styling(latex_options = c("scale_down", "hold_position", "centering"))
```

Now let's identify what those thresholds are for our model.
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
p <- 4
n <- nrow(rb_data_2022)

hat <- (2*p)/n
dffits_val <- 2*sqrt(p/n)
dfbetas_vals <- 2/sqrt(n)
covratio_val1 <- 1 - (3*p)/n
covratio_val2 <- 1 + (3*p)/n
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
results_df$`Studentized.Residuals` <- as.numeric(results_df$`Studentized.Residuals`)
results_df$`Hat.Values` <- as.numeric(results_df$`Hat.Values`)
results_df$`Cooks.D.Values` <- as.numeric(results_df$`Cooks.D.Values`)
results_df$`DFFITS.Values` <- as.numeric(results_df$`DFFITS.Values`)
results_df$`DFBETA.Values.rushing_td` <- as.numeric(results_df$`DFBETA.Values.rushing_td`)
results_df$`DFBETA.Values.fpts.g` <- as.numeric(results_df$`DFBETA.Values.fpts.g`)
results_df$`DFBETA.Values.rushing_yds` <- as.numeric(results_df$`DFBETA.Values.rushing_yds`)
results_df$`COVRATIO.Values` <- as.numeric(results_df$`COVRATIO.Values`)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
outlier_rows <- list(
  "Studentized Residuals" = which(abs(results_df$`Studentized.Residuals`) > 3),
  "Hat Values" = which(results_df$`Hat.Values` > round(hat, 4)),
  "Cook’s Distance" = which(results_df$`Cooks.D.Values` > 1),
  "DFFITS" = which(abs(results_df$`DFFITS.Values`) > round(dffits_val, 4)),
  "DFBETAs" = which(abs(results_df$`DFBETA.Values.rushing_td`) > round(dfbetas_vals, 4) |
                      abs(results_df$`DFBETA.Values.rushing_yds`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.fpts.g`) > round(dfbetas_vals, 4)),
  "COVRATIO" = which(results_df$`COVRATIO.Values` < round(covratio_val1,4) | 
                     results_df$`COVRATIO.Values` > round(covratio_val2,4))
)

# Display rows for each metric
print(outlier_rows)
```
Since DFBETAS and COVRATIO don't tell us much on their own, we'll only look at rows that have at least one other irregular flag.
```{r}
studentized_residuals <- c(63)
hat_values <- c(9,10,17,41)
cooks_distance <- integer(0)  # Empty
dffits <- c(6,7,63)
dfbetas <- c(6,7,9,13,16,21,24,41,50,63)

covratio <- c(1,2,3,6,10,17,41,61,63)
```
```{r}
all_indices <- sort(unique(c(studentized_residuals, hat_values, dffits)))
dfbetas_filtered <- dfbetas[dfbetas %in% all_indices]
covratio_filtered <- covratio[covratio %in% all_indices]
```
```{r}
df <- data.frame(
  Index = all_indices,
  Studentized_Residuals = ifelse(all_indices %in% studentized_residuals, "✓", ""),
  Hat_Values = ifelse(all_indices %in% hat_values, "✓", ""),
  DFFITS = ifelse(all_indices %in% dffits, "✓", ""),
  DFBETAs = ifelse(all_indices %in% dfbetas_filtered, "✓", ""),
  COVRATIO = ifelse(all_indices %in% covratio_filtered, "✓", "")
)

print(df)
```
```{r}
df %>%
  kable(format = "pipe", align = "c") %>%
  kable_styling(full_width = FALSE)
```
only 3 with 3 or more

Model B
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
rstudent_residuals = rstudent(model_B)
hat_values = hatvalues(model_B)
cooks_d = cooks.distance(model_B)
dfbetas_values <- as.data.frame(dfbetas(model_B))
dffits = dffits(model_B)
covratio = covratio(model_B)
```
```{r, echo=FALSE, warning=FALSE}
results_df <- data.frame(
  "RB" = rb_data_2022$name,
  "Studentized Residuals" = rstudent_residuals,
  "Hat Values" = hat_values,
  "Cooks D Values" = cooks_d,
  "DFBETA Values" = dfbetas_values,
  "DFFITS Values" = dffits,
  "COVRATIO Values" = covratio
)

results_df <- results_df %>%
  mutate(across(where(is.numeric), round, digits = 4))


kable(results_df, caption = "Model A: Regression Diagnostics Table", align="c", booktabs = TRUE) %>%
  kable_styling(latex_options = c("scale_down", "hold_position", "centering"))
```

Now let's identify what those thresholds are for our model.
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
p <- 5
n <- nrow(rb_data_2022)

hat <- (2*p)/n
dffits_val <- 2*sqrt(p/n)
dfbetas_vals <- 2/sqrt(n)
covratio_val1 <- 1 - (3*p)/n
covratio_val2 <- 1 + (3*p)/n
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
results_df$`Studentized.Residuals` <- as.numeric(results_df$`Studentized.Residuals`)
results_df$`Hat.Values` <- as.numeric(results_df$`Hat.Values`)
results_df$`Cooks.D.Values` <- as.numeric(results_df$`Cooks.D.Values`)
results_df$`DFFITS.Values` <- as.numeric(results_df$`DFFITS.Values`)
results_df$`DFBETA.Values.fpts.g` <- as.numeric(results_df$`DFBETA.Values.fpts.g`)
results_df$`DFBETA.Values.gp` <- as.numeric(results_df$`DFBETA.Values.gp`)
results_df$`DFBETA.Values.receving_td` <- as.numeric(results_df$`DFBETA.Values.receiving_td`)
results_df$`DFBETA.Values.rushing_td` <- as.numeric(results_df$`DFBETA.Values.rushing_td`)
results_df$`COVRATIO.Values` <- as.numeric(results_df$`COVRATIO.Values`)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
outlier_rows <- list(
  "Studentized Residuals" = which(abs(results_df$`Studentized.Residuals`) > 3),
  "Hat Values" = which(results_df$`Hat.Values` > round(hat, 4)),
  "Cook’s Distance" = which(results_df$`Cooks.D.Values` > 1),
  "DFFITS" = which(abs(results_df$`DFFITS.Values`) > round(dffits_val, 4)),
  "DFBETAs" = which(abs(results_df$`DFBETA.Values.fpts.g`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.receiving_td`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.gp`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.rushing_td`) > round(dfbetas_vals, 4)),
  "COVRATIO" = which(results_df$`COVRATIO.Values` < round(covratio_val1,4) | 
                     results_df$`COVRATIO.Values` > round(covratio_val2,4))
)

# Display rows for each metric
print(outlier_rows)
```
Since DFBETAS and COVRATIO don't tell us much on their own, we'll only look at rows that have at least one other irregular flag.
```{r}
studentized_residuals <- c(63)
hat_values <- c(10,17,41,50,61,64)
cooks_distance <- integer(0)  # Empty
dffits <- c(7,50,63)

dfbetas <- c(7,16,21,44,50,54,63)

covratio <- c(10,17,41,50,61,63,64)
```
```{r}
all_indices <- sort(unique(c(studentized_residuals, hat_values, dffits)))
dfbetas_filtered <- dfbetas[dfbetas %in% all_indices]
covratio_filtered <- covratio[covratio %in% all_indices]
```
```{r}
df <- data.frame(
  Index = all_indices,
  Studentized_Residuals = ifelse(all_indices %in% studentized_residuals, "✓", ""),
  Hat_Values = ifelse(all_indices %in% hat_values, "✓", ""),
  DFFITS = ifelse(all_indices %in% dffits, "✓", ""),
  DFBETAs = ifelse(all_indices %in% dfbetas_filtered, "✓", ""),
  COVRATIO = ifelse(all_indices %in% covratio_filtered, "✓", "")
)

print(df)
```
```{r}
df %>%
  kable(format = "pipe", align = "c") %>%
  kable_styling(full_width = FALSE)
```
2 with 3 or more

Model C
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
rstudent_residuals = rstudent(model_C)
hat_values = hatvalues(model_C)
cooks_d = cooks.distance(model_C)
dfbetas_values <- as.data.frame(dfbetas(model_C))
dffits = dffits(model_C)
covratio = covratio(model_C)
```
```{r, echo=FALSE, warning=FALSE}
results_df <- data.frame(
  "RB" = rb_data_2022$name,
  "Studentized Residuals" = rstudent_residuals,
  "Hat Values" = hat_values,
  "Cooks D Values" = cooks_d,
  "DFBETA Values" = dfbetas_values,
  "DFFITS Values" = dffits,
  "COVRATIO Values" = covratio
)

results_df <- results_df %>%
  mutate(across(where(is.numeric), round, digits = 4))


kable(results_df, caption = "Model A: Regression Diagnostics Table", align="c", booktabs = TRUE) %>%
  kable_styling(latex_options = c("scale_down", "hold_position", "centering"))
```

Now let's identify what those thresholds are for our model.
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
p <- 4
n <- nrow(rb_data_2022)

hat <- (2*p)/n
dffits_val <- 2*sqrt(p/n)
dfbetas_vals <- 2/sqrt(n)
covratio_val1 <- 1 - (3*p)/n
covratio_val2 <- 1 + (3*p)/n
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
results_df$`Studentized.Residuals` <- as.numeric(results_df$`Studentized.Residuals`)
results_df$`Hat.Values` <- as.numeric(results_df$`Hat.Values`)
results_df$`Cooks.D.Values` <- as.numeric(results_df$`Cooks.D.Values`)
results_df$`DFFITS.Values` <- as.numeric(results_df$`DFFITS.Values`)
results_df$`DFBETA.Values.rushing_td` <- as.numeric(results_df$`DFBETA.Values.rushing_td`)
results_df$`DFBETA.Values.rushing_att` <- as.numeric(results_df$`DFBETA.Values.rushing_att`)
results_df$`DFBETA.Values.fpts.g` <- as.numeric(results_df$`DFBETA.Values.fpts.g`)
results_df$`COVRATIO.Values` <- as.numeric(results_df$`COVRATIO.Values`)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
outlier_rows <- list(
  "Studentized Residuals" = which(abs(results_df$`Studentized.Residuals`) > 3),
  "Hat Values" = which(results_df$`Hat.Values` > round(hat, 4)),
  "Cook’s Distance" = which(results_df$`Cooks.D.Values` > 1),
  "DFFITS" = which(abs(results_df$`DFFITS.Values`) > round(dffits_val, 4)),
  "DFBETAs" = which(
                    abs(results_df$`DFBETA.Values.rushing_td`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.rushing_att`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.fpts.g`) > round(dfbetas_vals, 4)),
  "COVRATIO" = which(results_df$`COVRATIO.Values` < round(covratio_val1,4) | 
                     results_df$`COVRATIO.Values` > round(covratio_val2,4))
)

# Display rows for each metric
print(outlier_rows)
```
Since DFBETAS and COVRATIO don't tell us much on their own, we'll only look at rows that have at least one other irregular flag.
```{r}
studentized_residuals <- c(63)
hat_values <- c(10,17,41)
cooks_distance <- integer(0)  # Empty
dffits <- c(6,7,63)

dfbetas <- c(6,7,9,16,41,50,63)

covratio <- c(2,6,10,17,19,41,63)
```
```{r}
all_indices <- sort(unique(c(studentized_residuals, hat_values, dffits)))
dfbetas_filtered <- dfbetas[dfbetas %in% all_indices]
covratio_filtered <- covratio[covratio %in% all_indices]
```
```{r}
df <- data.frame(
  Index = all_indices,
  Studentized_Residuals = ifelse(all_indices %in% studentized_residuals, "✓", ""),
  Hat_Values = ifelse(all_indices %in% hat_values, "✓", ""),
  DFFITS = ifelse(all_indices %in% dffits, "✓", ""),
  DFBETAs = ifelse(all_indices %in% dfbetas_filtered, "✓", ""),
  COVRATIO = ifelse(all_indices %in% covratio_filtered, "✓", "")
)

print(df)
```
```{r}
df %>%
  kable(format = "pipe", align = "c") %>%
  kable_styling(full_width = FALSE)
```
3 with 3 or more

Model D
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
rstudent_residuals = rstudent(model_D)
hat_values = hatvalues(model_D)
cooks_d = cooks.distance(model_D)
dfbetas_values <- as.data.frame(dfbetas(model_D))
dffits = dffits(model_D)
covratio = covratio(model_D)
```
```{r, echo=FALSE, warning=FALSE}
results_df <- data.frame(
  "RB" = rb_data_2022$name,
  "Studentized Residuals" = rstudent_residuals,
  "Hat Values" = hat_values,
  "Cooks D Values" = cooks_d,
  "DFBETA Values" = dfbetas_values,
  "DFFITS Values" = dffits,
  "COVRATIO Values" = covratio
)

results_df <- results_df %>%
  mutate(across(where(is.numeric), round, digits = 4))


kable(results_df, caption = "Model A: Regression Diagnostics Table", align="c", booktabs = TRUE) %>%
  kable_styling(latex_options = c("scale_down", "hold_position", "centering"))
```

Now let's identify what those thresholds are for our model.
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
p <- 6
n <- nrow(rb_data_2022)

hat <- (2*p)/n
dffits_val <- 2*sqrt(p/n)
dfbetas_vals <- 2/sqrt(n)
covratio_val1 <- 1 - (3*p)/n
covratio_val2 <- 1 + (3*p)/n
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
results_df$`Studentized.Residuals` <- as.numeric(results_df$`Studentized.Residuals`)
results_df$`Hat.Values` <- as.numeric(results_df$`Hat.Values`)
results_df$`Cooks.D.Values` <- as.numeric(results_df$`Cooks.D.Values`)
results_df$`DFFITS.Values` <- as.numeric(results_df$`DFFITS.Values`)
results_df$`DFBETA.Values.receiving_td` <- as.numeric(results_df$`DFBETA.Values.receiving_td`)
results_df$`DFBETA.Values.receiving_yds` <- as.numeric(results_df$`DFBETA.Values.receiving_yds`)
results_df$`DFBETA.Values.rushing_td` <- as.numeric(results_df$`DFBETA.Values.rushing_td`)
results_df$`DFBETA.Values.rushing_yds` <- as.numeric(results_df$`DFBETA.Values.rushing_yds`)
results_df$`DFBETA.Values.fpts.g` <- as.numeric(results_df$`DFBETA.Values.fpts.g`)
results_df$`COVRATIO.Values` <- as.numeric(results_df$`COVRATIO.Values`)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
outlier_rows <- list(
  "Studentized Residuals" = which(abs(results_df$`Studentized.Residuals`) > 3),
  "Hat Values" = which(results_df$`Hat.Values` > round(hat, 4)),
  "Cook’s Distance" = which(results_df$`Cooks.D.Values` > 1),
  "DFFITS" = which(abs(results_df$`DFFITS.Values`) > round(dffits_val, 4)),
  "DFBETAs" = which(
                    abs(results_df$`DFBETA.Values.receiving_td`) > round(dfbetas_vals, 4) |
                    
                    abs(results_df$`DFBETA.Values.receiving_yds`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.rushing_td`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.rushing_yds`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.fpts.g`) > round(dfbetas_vals, 4)),
  "COVRATIO" = which(results_df$`COVRATIO.Values` < round(covratio_val1,4) | 
                     results_df$`COVRATIO.Values` > round(covratio_val2,4))
)

# Display rows for each metric
print(outlier_rows)
```
Since DFBETAS and COVRATIO don't tell us much on their own, we'll only look at rows that have at least one other irregular flag.
```{r}
studentized_residuals <- c(63)
hat_values <- c(1,2,7,9,10,17,19,21,41,50,54,61)
cooks_distance <- integer(0)  # Empty
dffits <- c(6,7,41,50,63)

dfbetas <- c(6,7,16,20,21,41,45,50,54,63)

covratio <- c(1,  2,  3,  4,  5,  6,  8,  9, 10, 14, 17, 19, 26, 41, 46, 49, 50, 58, 61, 63, 67)
```
```{r}
all_indices <- sort(unique(c(studentized_residuals, hat_values, dffits)))
dfbetas_filtered <- dfbetas[dfbetas %in% all_indices]
covratio_filtered <- covratio[covratio %in% all_indices]
```
```{r}
df <- data.frame(
  Index = all_indices,
  Studentized_Residuals = ifelse(all_indices %in% studentized_residuals, "✓", ""),
  Hat_Values = ifelse(all_indices %in% hat_values, "✓", ""),
  DFFITS = ifelse(all_indices %in% dffits, "✓", ""),
  DFBETAs = ifelse(all_indices %in% dfbetas_filtered, "✓", ""),
  COVRATIO = ifelse(all_indices %in% covratio_filtered, "✓", "")
)

print(df)
```
```{r}
df %>%
  kable(format = "pipe", align = "c") %>%
  kable_styling(full_width = FALSE)
```
5 in three or more

Model E
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
rstudent_residuals = rstudent(model_E)
hat_values = hatvalues(model_E)
cooks_d = cooks.distance(model_E)
dfbetas_values <- as.data.frame(dfbetas(model_E))
dffits = dffits(model_E)
covratio = covratio(model_E)
```
```{r, echo=FALSE, warning=FALSE}
results_df <- data.frame(
  "RB" = rb_data_2022$name,
  "Studentized Residuals" = rstudent_residuals,
  "Hat Values" = hat_values,
  "Cooks D Values" = cooks_d,
  "DFBETA Values" = dfbetas_values,
  "DFFITS Values" = dffits,
  "COVRATIO Values" = covratio
)

results_df <- results_df %>%
  mutate(across(where(is.numeric), round, digits = 4))


kable(results_df, caption = "Model A: Regression Diagnostics Table", align="c", booktabs = TRUE) %>%
  kable_styling(latex_options = c("scale_down", "hold_position", "centering"))
```

Now let's identify what those thresholds are for our model.
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
p <- 7
n <- nrow(rb_data_2022)

hat <- (2*p)/n
dffits_val <- 2*sqrt(p/n)
dfbetas_vals <- 2/sqrt(n)
covratio_val1 <- 1 - (3*p)/n
covratio_val2 <- 1 + (3*p)/n
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
results_df$`Studentized.Residuals` <- as.numeric(results_df$`Studentized.Residuals`)
results_df$`Hat.Values` <- as.numeric(results_df$`Hat.Values`)
results_df$`Cooks.D.Values` <- as.numeric(results_df$`Cooks.D.Values`)
results_df$`DFFITS.Values` <- as.numeric(results_df$`DFFITS.Values`)
results_df$`DFBETA.Values.fpts.g` <- as.numeric(results_df$`DFBETA.Values.fpts.g`)
results_df$`DFBETA.Values.gp` <- as.numeric(results_df$`DFBETA.Values.gp`)
results_df$`DFBETA.Values.receiving_td` <- as.numeric(results_df$`DFBETA.Values.receiving_td`)
results_df$`DFBETA.Values.receiving_tgts` <- as.numeric(results_df$`DFBETA.Values.receiving_tgts`)
results_df$`DFBETA.Values.receiving_yds` <- as.numeric(results_df$`DFBETA.Values.receiving_yds`)
results_df$`DFBETA.Values.rushing_td` <- as.numeric(results_df$`DFBETA.Values.rushing_td`)
results_df$`COVRATIO.Values` <- as.numeric(results_df$`COVRATIO.Values`)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
outlier_rows <- list(
  "Studentized Residuals" = which(abs(results_df$`Studentized.Residuals`) > 3),
  "Hat Values" = which(results_df$`Hat.Values` > round(hat, 4)),
  "Cook’s Distance" = which(results_df$`Cooks.D.Values` > 1),
  "DFFITS" = which(abs(results_df$`DFFITS.Values`) > round(dffits_val, 4)),
  "DFBETAs" = which(abs(results_df$`DFBETA.Values.fpts.g`) > round(dfbetas_vals, 4) |
                      abs(results_df$`DFBETA.Values.gp`) > round(dfbetas_vals, 4) |
                      abs(results_df$`DFBETA.Values.receiving_td`) > round(dfbetas_vals, 4) |
                      abs(results_df$`DFBETA.Values.receiving_tgts`) > round(dfbetas_vals, 4) |
                      abs(results_df$`DFBETA.Values.receiving_yds`) > round(dfbetas_vals, 4) |
                    abs(results_df$`DFBETA.Values.rushing_td`) > round(dfbetas_vals, 4)),
  "COVRATIO" = which(results_df$`COVRATIO.Values` < round(covratio_val1,4) | 
                     results_df$`COVRATIO.Values` > round(covratio_val2,4))
)

# Display rows for each metric
print(outlier_rows)
```
Since DFBETAS and COVRATIO don't tell us much on their own, we'll only look at rows that have at least one other irregular flag.
```{r}
studentized_residuals <- c(63)
hat_values <- c(2,7,10,17,50,64)
cooks_distance <- integer(0)  # Empty
dffits <- c(7,50,63)
dfbetas <- c(2, 7, 13, 16, 20, 21, 24, 45, 48, 50, 63, 65)

covratio <- c(2,  9, 10, 12, 17, 41, 50, 61, 63, 64)
```
```{r}
all_indices <- sort(unique(c(studentized_residuals, hat_values, dffits)))
dfbetas_filtered <- dfbetas[dfbetas %in% all_indices]
covratio_filtered <- covratio[covratio %in% all_indices]
```
```{r}
df <- data.frame(
  Index = all_indices,
  Studentized_Residuals = ifelse(all_indices %in% studentized_residuals, "✓", ""),
  Hat_Values = ifelse(all_indices %in% hat_values, "✓", ""),
  DFFITS = ifelse(all_indices %in% dffits, "✓", ""),
  DFBETAs = ifelse(all_indices %in% dfbetas_filtered, "✓", ""),
  COVRATIO = ifelse(all_indices %in% covratio_filtered, "✓", "")
)

print(df)
```
```{r}
df %>%
  kable(format = "pipe", align = "c") %>%
  kable_styling(full_width = FALSE)
```
only 4 with 3 or more

Accuracy Checking
unscale
```{r}
undo_standardization <- function(df, original_train_data) {
  means <- attr(original_train_data, "means") %>% as.list() %>% unlist()  # Convert to named vector
  sds <- attr(original_train_data, "sds") %>% as.list() %>% unlist()  # Convert to named vector

  numeric_cols <- df %>% dplyr::select(where(is.numeric)) %>% names()  # Select only numeric columns

  df %>%
    mutate(across(all_of(numeric_cols), ~ . * sds[cur_column()] + means[cur_column()]))
}

rb_data_2022 <- undo_standardization(rb_data_2022, rb_data_2022)
```

report mallows, press, adj r sq, number of problem children in 3 or more

# Model Validation - k fold cross validation
```{r}
k_folds = 5

#set up cross fold
set.seed(24)
folds <- createFolds(rb_data_2022$`2023_fpts/g`, k = k_folds, list = TRUE)

#intialize results df
cv_results <- data.frame(Model = character(), RMSE = numeric(), R2 = numeric(), PRESS = numeric(), R2_pred = numeric(), stringsAsFactors = FALSE)

#establish candidate models
candidate_models <- list(ModelA = `2023_fpts/g` ~ fpts.g + rushing_td+rushing_yds,
                         ModelB = `2023_fpts/g` ~ fpts.g +gp+receiving_td+rushing_td,
                         ModelC = `2023_fpts/g` ~ fpts.g + rushing_att+rushing_td,
                         ModelD = `2023_fpts/g` ~ fpts.g + receiving_td+receiving_yds+rushing_td+rushing_yds,
                         ModelE = `2023_fpts/g` ~ fpts.g+gp+receiving_td+receiving_tgts+receiving_yds+rushing_td)


#loop through
for (i in seq_along(candidate_models)) {
  
  #model formula
  model_formula <- candidate_models[[i]]
  
  #result vectors
  rmse_values <- c()
  r2_values <- c()
  press_values <- c()

  for (j in seq_along(folds)) {
    #split
    test_indices <- folds[[j]]
    train_data <- rb_data_2022[-test_indices, ]
    test_data <- rb_data_2022[test_indices, ]
    
    #only numeric
    numeric_cols <- train_data %>%
      dplyr::select(where(is.numeric)) %>%
      dplyr::select(-`2023_fpts/g`) %>%  # Exclude target variable
      colnames()
    
    #standardize train
    train_mean <- train_data %>%
      summarise(across(all_of(numeric_cols), mean, na.rm = TRUE))
    
    train_sd <- train_data %>%
      summarise(across(all_of(numeric_cols), sd, na.rm = TRUE))
    
    train_data_std <- train_data %>%
      mutate(across(all_of(numeric_cols), ~ (. - train_mean[[cur_column()]]) / train_sd[[cur_column()]]))
    
    #standardize test
    test_mean <- test_data %>%
      summarise(across(all_of(numeric_cols), mean, na.rm = TRUE))
    
    test_sd <- test_data %>%
      summarise(across(all_of(numeric_cols), sd, na.rm = TRUE))
    
    test_data_std <- test_data %>%
      mutate(across(all_of(numeric_cols), ~ (. - test_mean[[cur_column()]]) / test_sd[[cur_column()]]))
    
    #train model
    model <- lm(model_formula, data = train_data_std)
    
    #predict
    predictions_std <- predict(model, test_data_std)

    #unstandardize
    y_mean <- mean(test_data$`2023_fpts/g`)
    y_sd <- sd(test_data$`2023_fpts/g`)
    
    predictions <- (predictions_std * y_sd) + y_mean  # Reverse transformation
    
    #compute RMSE and R^2
    rmse <- sqrt(mean((test_data$`2023_fpts/g` - predictions)^2))
    r2 <- cor(test_data$`2023_fpts/g`, predictions)^2
    
    #compute PRESS
    press_values <- compute_press(model)
    
    #store values
    rmse_values <- c(rmse_values, rmse)
    r2_values <- c(r2_values, r2)
  }
  
  #compute SST (Total Sum of Squares) for entire dataset
  SST <- sum((rb_data_2022$`2023_fpts/g` - mean(rb_data_2022$`2023_fpts/g`))^2)
  
  #compute mean PRESS and R2_pred
  mean_PRESS <- mean(press_values)
  R2_pred <- 1 - (mean_PRESS / SST)
  
  #store mean cross-validated RMSE, R2, PRESS, and R2_pred for this model
  cv_results <- rbind(cv_results, data.frame(
    Model = as.character(deparse(model_formula)),  # Convert formula properly
    RMSE = mean(rmse_values),
    R2 = mean(r2_values),
    PRESS = mean_PRESS,
    R2_pred = R2_pred
))

}

#sort models by rmse
cv_results <- cv_results[order(cv_results$PRESS), ]

print(cv_results)
```
```{r}
coef(model_A)
coef(model_B)
coef(model_C)
coef(model_D)
coef(model_E)
```

# Predictions
Apply to 2023 data and save predictions and rankings
selecting model A
```{r}
rb_data_2023 <- read.csv("rb_data_2023.csv")
```
```{r}
rb_data_2023$predicted_fpts_2024 <- predict(model_B, newdata = rb_data_2023)
print(rb_data_2023$predicted_fpts_2024)
```
```{r}
rb_predictions <- rb_data_2023 %>%
  dplyr::select(name, `X2024_fpts.g`, predicted_fpts_2024) %>%
  mutate(position="RB")

head(rb_predictions)
```
Check accuracy for fun??
```{r}
write.csv(rb_predictions, "rb_predictions.csv", row.names = FALSE)
```